---
title: "Taxonomy Quick View"
author: Suggestions to imetalabca@gmail.com
date: Report generated @`r Sys.time()`
always_allow_html: TRUE
output:
  html_document:
    fig_width: 10
    fig_caption: TRUE
    toc: TRUE
    toc_depth: 4
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
    number_sections: TRUE
  pdf_document:
    toc: TRUE
    toc_depth: 4
    highlight: tango
    df_print: kable
    number_sections: TRUE
  word_document:
    toc: TRUE
    toc_depth: 4
    highlight: tango
    df_print: kable
    smart: TRUE
params:  
  data_table: !r NULL ### Manditory; main data input for the report
  meta_table: !r NULL ### Optional; If provided, will do some statistics 
  external_message: !NULL
  if_html: TRUE # if you want to pdf or word export, set this as FALSE
---

<style type="text/css">
.main-container {
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r changelog, echo=FALSE, eval= FALSE}

# 20220203 change the input  parameter from input_datatable to data_table, to be unified as the one function
# 20210804 added percent stacked bar plots
# 20210608 added Meta information; added downloadable tables
# 20190819


# the input file is tree/lineage file, with all lineage information on different taxon levels, followed by intensity columns across experiments
# required columns in input file (taxon)

  # 1: Name
  # 2: Rank
  # 3: Superkingdom
  # 4: Kingdom
  # 5: Phylum
  # 6: Class
  # 7: Order
  # 8: Family
  # 9: Genus
  # 10: Species
  # Intensity a
  # Intensity b
  # ... more columns of intensity


```

```{r echo = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE, cache = FALSE, error = TRUE)
```

```{r setup}

htmltools::img(src = system.file("rmd/iMetaReport.png", package = "metareport"), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;width:100px;height:100px;')
library(tidyverse)
library(ggplot2)
library(plotly)
library(readxl)
#library(pheatmap)
library(reshape2)
library(vegan)
#library(ggdendro)
library(gridExtra)
library(DT)# for interactive data table display 
library(rmetalab)
library(d3r)
library(sunburstR)
library(d3Tree)
```




```{r file_taxonomy, echo=FALSE, fig.width= 15,fig.height=10}


data_table <- params$data_table
meta_table <- params$meta_table


taxon_summary <- tidy_taxon_table(data_table)
experiment_names <- taxon_summary$experiment

# check meta_table
if( is.null(meta_table)){ 
  meta_info <- "* **No meta information provided**"
  
}else if (any(is.na(meta_table))){ 
  
  meta_info <-  "* **The meta information provided has missing values, please check, analysis continues without grouping information**"
  meta_table <- NULL
  
}else if(!identical(sort(toupper(experiment_names)),sort(toupper(meta_table[,2])))){
  meta_info <-  "* **The experiment setting in meta table provided do not match the experiment in the data_table, please check, analysis continues without grouping information**"
  meta_table <- NULL

}else{
  meta_table <- meta_table[,-1] # remove the raw files column, not needed in this report
  meta_table <- meta_table[match(toupper(experiment_names),toupper(meta_table[,1])),] # ensure the order is the same is the column names
  meta_info <- NULL
  # numerate all meta info
  for(i in 2: ncol(meta_table)){
     meta_info <-   paste0(meta_info, "* **Grouping ",i-1,": **" , paste0(unique(meta_table[,i]), collapse="; "), "<br>  \n\n" )  
  }
}



```


```{r error, results='asis'}
# use this as a template for conditional multi-line in-line code
if(!is.null(params$external_message)){
  cat("# Something went wrong\n")
  cat(" **In most cases when you see this message, the report is either not complete or based on incomplete importation of the data table** \n")
  cat(" **Please doublecheck the input file or the server status according to the message below** \n")
  cat(params$external_message)
} 

```



# Intro

**This report provides some basic description and visualization of the MetaLab taxonomy results. **
**The report is based on the taxa.refine.txt, which is a node based hierarchical like data**
**Users can use this to quickly check the taxonomic profile of the dataset at each taxonomic level.**
**For more taxon tree view, please go to MetaMap report section **


# Sample overview

*  **Number of samples in your dataset: ** `r taxon_summary$experiment_number `
*  **Number of species identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$species)`
*  **Number of genera identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$genus)`
*  **Number of families identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$family)`
*  **Number of orders identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$order)`
*  **Number of classes identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$class)`
*  **Number of phyla identified: ** `r nrow(taxon_summary$taxon_summary_widetable_list$phylum)`
`r meta_info` 


```{r include=FALSE}
# the purpose of this block is only to inlucde plotly, then ggplotly will be able to display in the for loop:
# The only working way is to  create a plot outside the loop.
# see reference: https://github.com/plotly/plotly.R/issues/273
ggplotly(qplot(1, 1))
```



# Overall pattern

<details>
  <summary>About this plot</summary>
  
  * Intensity is based on the total intensity across samples
  * Graph is plotted by suburstR::sund2b
  * Tree is plotted by networkD3::diagonalNetwork
    * Node without labeling is not present

</details>

## sunburst

```{r fig.width= 10, fig.height= 10}

df_tree_sunburst <- data_table %>% select(Superkingdom,Phylum, Class,Order,Family,Genus,Species) %>% na_if("") %>% 
  cbind(.,intensity = rowSums(data_table[,11:ncol(data_table)]))

sund2b(d3_nest(df_tree_sunburst,value_cols = "intensity"),
       valueField = "intensity",
       showLabels =  TRUE)

# 
# sunburst(
#   d3_nest(df_tree,value_cols = "intensity"),
#   valueField = "intensity",
#   sumNodes = FALSE,
#   withD3 = FALSE
# )
## to do

fig.height_tree <-   nrow(df_tree_sunburst)*0.03 + 6

```

## Tree

```{r fig.width= 12, fig.height= fig.height_tree}

tree <- df2tree(struct = df_tree_sunburst,rootname = 'root')
networkD3::diagonalNetwork(tree)
# d3tree(list(root = df2tree(struct = df_tree_sunburst,rootname = 'root'),
#         layout = 'collapse')
#       )  
# d3tree(list(root = tree,
#         layout = 'cartesian')
#       )  
#networkD3::radialNetwork(tree)


```



# Species level Sample Clustering/Similarity

<details>
  <summary>About this clustering Diversity</summary>
  
  * sample clustering based on the species-level composition
  * Distance measure used is "euclidean" (dist(x,method = "euclidean"))
  * clustering method: hclust(distance, method= "ward.D")
  * Clustering analysis will be performed when there are more than 2 samples

</details>


```{r}
fig.height_hca <- taxon_summary$experiment_number*0.2 +5
```


```{r cluster_fig, echo=FALSE,fig.width= 8, fig.height=fig.height_hca}
count_species <- taxon_summary$taxon_summary_widetable_list$species


if(ncol(count_species) >2){
  plot_HCA(count_species, meta_table = meta_table[,-1])
}else{
  print("There are not enough samples to do clustering.")
}
```





# Richness  {.tabset .tabset-fade }

<details>
  <summary>What is richness</summary>
  
  * Identification count on different levels
  * [ref
  * More detailed [References](https://cran.r-project.org/web/packages/tabula/vignettes/diversity.html)
  * R package used for calculation: tabula
  

</details>

```{r identification_fig, echo=FALSE, fig.width= 10, fig.height=5}
# plot on these levels ("Species", "Genus", "Family", "Order", "Class", "Phylum")
ID_number_table <- taxon_summary$ID_number_table
plot_list <- plot_taxon_table(ID_number_table, meta_table = meta_table)

```


```{r echo=FALSE, results='asis',error=TRUE, eval= TRUE, fig.width= 10, fig.height=8}

for(plot in names(plot_list)){
  cat(paste0("## ", plot,  "\n"))
  p = plot_list[[plot]]
  print(htmltools::tagList(ggplotly(p)))
  cat("\n***\n")
}

```


## Result table for download

```{r identification_data, echo=FALSE,fig.width= 10,fig.height=8, cache = FALSE}

if(!is.null(meta_table)){
  ID_number_table <- cbind(meta_table,taxon_summary$ID_number_table)
}

datatable(ID_number_table, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           scrollX = TRUE,
                           buttons = c('copy', 'csv', 'excel'),
                           pageLength = 10))

```



# Heterogeneity and Evenness {.tabset .tabset-fade }


<details>
  <summary>Alpha Diversity</summary>
  
  * Heterogeneity and Evenness belongs to the concept of alpha diversity
  * We only show Shannon-Wiener index in this plot
  * the diversity was calculated by vegan::diversity(x,index = "shannon")
  * more details about what is types of diversity and the way of calculation
  * [References](https://cran.r-project.org/web/packages/tabula/vignettes/diversity.html)
  * R package used: vegan and tabula
  

</details>

```{r alpha_fig, echo=FALSE}
alpha_diversity_table <- taxon_summary$alpha_diversity$index_diversity_shannon
plot_list <- plot_taxon_table(alpha_diversity_table, meta_table = meta_table, ylabel = "Alpha Diversity: Shanon")

```

```{r echo=FALSE, results='asis',error=TRUE, eval= TRUE, fig.width= 10, fig.height=8}

for(plot in names(plot_list)){
  cat(paste0("## ", plot,  "\n"))
  p = plot_list[[plot]]
  print(htmltools::tagList(ggplotly(p)))
  cat("\n***\n")
}

```



## Result table for download

```{r alpha_data, echo=FALSE,fig.width= 8,fig.height=4, cache = FALSE}
if(!is.null(meta_table)){
  alpha_diversity_table <- cbind(meta_table,alpha_diversity_table)
}

datatable(alpha_diversity_table, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           scrollX = TRUE,
                           buttons = c('copy', 'csv', 'excel'),
                           pageLength = 10))


```


# Beta diversity {.tabset .tabset-fade }

<details>
  <summary>Beta Diversity</summary>
  
  * here we only show Bray Curtis dissimilarity as one of the beta diversity
  * The data is based on species level
  * The caclculation is done by vegan::vegdist(t(count_species),method="bray")
  * Visualization by PCoA
  * more details about what is types of diversity and the way of calculation
  * References: https://en.wikipedia.org/wiki/Bray%E2%80%93Curtis_dissimilarity
  * R package used: vegan and tabula
  
</details>

```{r beta-diversity calculation}

count_species <- taxon_summary$taxon_summary_widetable_list$species
beta_diversity <- plot_beta_diversity(count_species =count_species, meta_table = meta_table)

```

```{r echo=FALSE, results='asis',error=TRUE, eval= TRUE, fig.width= 10, fig.height=8}

for(plot in names(beta_diversity$plot_list)){
  cat(paste0("## ", plot,  "\n"))
  p = beta_diversity$plot_list[[plot]]
  print(htmltools::tagList(ggplotly(p)))
  cat("\n\n")
}

```

## Result table for download

```{r beta_data, echo=FALSE,fig.width= 8,fig.height=4, cache = FALSE}
datatable(beta_diversity$PCoA.bray, extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           scrollX = TRUE,
                           buttons = c('copy', 'csv', 'excel'),
                           pageLength = 10))


```





# Stacked barplots based on biomass {.tabset .tabset-fade }


```{r echo=FALSE, results='asis',error=TRUE, eval= TRUE, fig.width= 10, fig.height=8}

tab_list <- taxon_summary$taxon_summary_longtable_list

for(level in names(tab_list)){

  cat(paste0("## ", level,  "\n"))

  p2 <- ggplot(tab_list[[level]], aes(x = Sample, y = Intensity, fill = Name)) +
       geom_bar(stat = "identity") + theme_bw() +
       ylab("Intensity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)
  )
  print(htmltools::tagList(ggplotly(p2)))
  
  
  cat("\n\n")
}



```



# Normalized bar-plots based on biomass {.tabset .tabset-fade }

```{r echo=FALSE, results='asis',error=TRUE, eval= TRUE, fig.width= 10, fig.height=8}

tab_list <- taxon_summary$taxon_summary_longtable_list

for(level in names(tab_list)){

  cat(paste0("## ", level,  "\n"))

  p2 <- ggplot(tab_list[[level]], aes(x = Sample, y = Intensity, fill = Name)) +
       geom_col(position = "fill") + theme_bw() +
       ylab("Intensity") + theme(axis.text.x = element_text(angle = 45, hjust = 1)
  )
  print(htmltools::tagList(ggplotly(p2)))
  cat("\n\n")
}

```

