
<!-- input: df_intensity, pure data.matrix with intensity -->
<!-- input optional: meta_data -->

```{r echo=FALSE, error=TRUE, results='asis'}
df_intensity_Q100 <- df_intensity[which(rowSums(df_intensity > 0) == ncol(df_intensity)) , , drop = FALSE] # the Q100 features/rows intensity matrix

# only use top 100 for
if (nrow(df_intensity_Q100) > 100){
  df_intensity_for_interactive_plot<- df_intensity_Q100 %>% slice_max(order_by = rowSums(.), n =100) %>% log10 %>%  as.matrix()
  #cat("* Only top100 items (total intensity) for interactive heatmap plotting")
}else if(nrow(df_intensity_Q100) == 0){
  df_intensity_for_interactive_plot <- (df_intensity+1) %>% slice_max(order_by = rowSums(.), n =100) %>%  log10 %>% as.matrix()
  cat("* No Q100 items, use top100 features (total intensity) for interactive heatmap plotting")
}else{
  df_intensity_for_interactive_plot <- df_intensity_Q100 %>% log10 %>%  as.matrix()
}

```

* There are `r nrow(df_intensity)` features.
* With `r nrow(df_intensity_Q100)` 100% presence across experiments (Q100).


## Interactive heatmap

<details>

  <summary>About this plot</summary>
  
  * If there are more than 100 Q100 features, only the top100 will be used for this interactive plot
  * if there are no Q100 features, top100 will be used with +1 
  * data is transformed by log10
  * Scale on feature wise to be mean as 0, and sd as 1
  * plot was generated by heatmaply::heatmaply

</details>


```{r echo=FALSE, fig.width= 8, fig.height=6, error=TRUE}

quickpheatmap(df_intensity_for_interactive_plot, scale = "row", column_meta = meta_table, interactive  = TRUE )

```

## Static Heatmap

<details>

  <summary>About this plot</summary>
  
  * All Q100 features are used for this static plot, if there are no Q100, Q50 will be used, by +1
  * transform by log10
  * Scale on feature wise to be mean as 0, and sd as 1
  * Plot was generated by complexheatmap::pheatmap

</details>


```{r echo=FALSE, fig.width= 8, fig.height=6, error=TRUE}

df_intensity_Q50 <- df_intensity[which(rowSums(df_intensity > 0) >= ncol(df_intensity)*0.5) , , drop = FALSE]
# # plot a heatmap of Q100
if(nrow(df_intensity_Q100) == 0 && nrow(df_intensity_Q50) > 0){
  quickpheatmap(df_intensity_Q50, scale = "row", column_meta = meta_table )
}else if( nrow(df_intensity_Q50) == 0){
  cat("Too many missing values, even 0 for Q50 , check the interactive plot of the top 100 features ")
}else{
  quickpheatmap(df_intensity_Q100, scale = "row", column_meta = meta_table )
}

```

## Hierarchical cluster

<details>

  <summary>About this plot</summary>
  
  * All intensity values +1
  * transform by log10
  * Scale on feature wise to be mean as 0, and sd as 1
  * Calculate the distance between features, using euclidean method
  * Do Hierarchical cluster analysis, with method = "complete"
  
</details>



```{r}
figure_height <- 0.1*ncol(df_intensity)+4

```

```{r echo=FALSE, fig.width= 8, fig.height = figure_height, error=TRUE}

plot_HCA(df_intensity = df_intensity, meta_table = meta_table[,-1,drop = FALSE])# delete the experiment column, only keep the grouping columns

```






