---
title: "Protein Profiling"
author: Suggestions to imetalabca@gmail.com
date: Report generated @`r Sys.time()`
output:
  html_document:
    fig_width: 10
    fig_caption: TRUE
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
    number_sections: true
params:  
  data_table: !r NULL ### Manditory; main data input for the report
  meta_table: !r NULL ### Optional; If provided, will do some statistics
  external_message: !NULL
  if_html: TRUE # if you want to pdf or word export, set this as FALSE
---

<style type="text/css">
.main-container {
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, echo=FALSE}
# change log
# 20220203 change the input  parameter from input_datable to data_table, to be unified as the one function



```




```{r,setup, echo=FALSE, warning = FALSE, message = FALSE}

htmltools::img(src = system.file("rmd/iMetaReport.png", package = "metareport"),
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:10px;width:100px;height:100px;')

# enviroment setup
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE, cache = FALSE, error = TRUE)

library(tidyverse)
library(ggplot2)
library(d3heatmap)
library(plotly)
library(rmetalab)
library(gplots)
library(rhandsontable)

```
```{r errorMEssage, results='asis'}
# use this as a template for conditional multi-line in-line code
if(!is.null(params$external_message)){
  cat("# Something went wrong\n")
  cat(" **In most cases when you see this message, the report is either not complete or based on incomplete importation of the data table** \n")
  cat(" **Please doublecheck the input file or the server status according to the message below** \n")
  cat(params$external_message)
} 

```


```{r collect_input, echo=FALSE,fig.width= 15,fig.height=10}
# collect input
data_table <- params$data_table
meta_table <- params$meta_table 

# process the data_table
data_table_tidied <- tidy_protein_table(data_table)  # tidy and process,
df_intensity <- data_table_tidied$intensity_matrix 


# Note: The folling analysis with meta info assumes that
# 1st columns as sample name, 2nd column as experiment name, 3rd column and after as grouping

# check meta_table
if( is.null(meta_table)){ 
  meta_info <- "* **No meta information provided**"
  
}else if (any(is.na(meta_table))){ 
  
  meta_info <-  "* **The meta information provided has missing values, please check, analysis continues without grouping information**"
  meta_table <- NULL
  
}else if(!identical(sort(data_table_tidied$experiment),sort(meta_table[,2]))){
  meta_info <-  "* **The experiment setting in meta table provided do not match the experiment in the data_table, please check, analysis continues without grouping information**"
  meta_table <- NULL

}else{
  meta_table <- meta_table[,-1] # remove the raw files column, not needed in this report
  meta_table <- meta_table[match(colnames(df_intensity),meta_table[,1]),] # ensure the order is the same is the column names
  meta_info <- NULL
  # numerate all meta info
  for(i in 2: ncol(meta_table)){
     meta_info <-   paste0(meta_info, "* **Grouping ",i-1,": **" , paste0(unique(meta_table[,i]), collapse="; "), "<br>  \n\n" )  
  }
}

```



# Intro

**This report provides some basic description of the protein(groups) identification. **

**Users can use this to quickly check the overal quality of the experiment**



# Quick Message


*  **Number of contaminant: ** `r data_table_tidied$n_contaminant`

*  **Number of reversed: ** `r data_table_tidied$n_reversed`

*  **Number of qualified proteingroups: ** `r ncol(df_intensity)`

*  **Number of experiment:  ** `r ncol(df_intensity)`

*  **All experiments:  ** `r colnames(df_intensity)`

`r meta_info` 


# Protein identification profiling  {.tabset .tabset-fade }


```{r, echo=FALSE, results='asis', error=TRUE, eval= TRUE}
res <- knitr::knit_child(system.file("rmd/ReportTemplate_child_protein_profile.Rmd", package = "metareport"), envir = environment(), quiet = TRUE)
cat(res, sep = '\n')

```



# Heatmap and Hierarchical clustering {.tabset .tabset-fade }

```{r, echo=FALSE, results='asis', error=TRUE, eval= TRUE}
if(ncol(df_intensity) > 1){ # heatamap, only if the quantification matrix has more than 1 columns 
  res <- knitr::knit_child(system.file("rmd/ReportTemplate_child_heatmap.Rmd", package = "metareport"),quiet = TRUE) # envir = environment(), 
  cat(res, sep = '\n')
}else{
  cat("There is only one column/experiment, no heatmap needed!")
}


```




# PCA Analysis {.tabset .tabset-fade }

```{r, echo=FALSE, results='asis', error=TRUE, eval= TRUE}

if(length(data_table_tidied$experiment ) > 3){
  
  if(nrow(data_table_tidied$intensity_matrix_Q100) ==0){
    cat("* No Q100 items, using top100 with intensity+1 for analyis")
  }
  
  res <- knitr::knit_child('ReportTemplate_child_PCA.Rmd', quiet = TRUE)
  cat(res, sep = '\n')
  
  # for extra plot with grouping information
  if(!is.null(meta_table)){
    for(i in 2: ncol(meta_table)){ 
       section_title <- paste0("Grouping: " , paste0(unique(meta_table[,i]), collapse="; ") )
       res <- knitr::knit_child(system.file("rmd/ReportTemplate_child_PCA_grouping.Rmd", package = "metareport"), quiet = TRUE)
       cat(res, sep = '\n')
   }
  }
  
}

```





```{r echo=FALSE,results='asis', error=TRUE, eval= TRUE}
  
  if(!is.null(meta_table)){ # only do this when meta provided
        cat("# Differential Analysis {.tabset .tabset-fade } \n")
    
    cat ("<details> \n")

    cat ("<summary>About this section</summary> \n")
    cat("* t-test will be performed for two samples setting, while annova will be performed for 3 and more samples \n")
    cat("* Note that this Differential Analysis is performed on only Q100 peptides for quick profiling \n")
    cat("* Only do analysis 1 with more than (including) two groups, 2: each group has replicates. \n")
    
    cat ("</details> \n")
    

    for(i in 2: ncol(meta_table)){ # i is the column index
      
       section_title<- paste0("Grouping: " , paste(unique(meta_table[,i]), collapse="; ") )
       
       if(length(unique(meta_table[,i])) >=2  && all(table(meta_table[,i]) >=2)){
         # will do ANOVA analysis block while: 
         # 1: more than (including) 3 groups, 2: each group has replicates.
            res <- knitr::knit_child(system.file("rmd/ReportTemplate_child_protein_DA.Rmd", package = "metareport"), quiet = TRUE)
            cat(res, sep = '\n')
       }
   }
  }
  


```


```{r echo=FALSE,fig.width= 8,fig.height=6, eval= FALSE}
# will do this block while: 1, more than (includeing) two meta/groups, 2, each group has more than (includeing) 2 samples.
if(length(unique(meta_table[,3])) >=2 && all(table(meta_table[,3]) >=2)){

pvalues_anova <- matrix_PostHoc(df_intensity,meta_table[,3])

df_intensity_p <- cbind(pvalues_anova, df_intensity)

rownames(df_intensity_p) <- protein_primary_ids

df_intensity_p_filtered <- df_intensity_p[which(!is.na(df_intensity_p$p_PostHoc_pairs)),] 

DT::datatable(df_intensity_p_filtered, extensions = 'Buttons',options = list(dom = "Blfrtip",scrollX=TRUE, buttons = list('copy', 'print', list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      ))))

}


```


