


## Intensity

<details>
  <summary>About this plot</summary>
  
  * A distribution of the density across samples/experiment
  * A direct and rough evidence to tell if needed to normalize across samples

</details>


```{r}
figure_height <- 0.1*ncol(df_intensity)+4

```


```{r protein distribution plot, echo=FALSE,fig.width= 8,fig.height=figure_height}

#note: in this log10 intenisty matrix, 0 is converted into infinity, therefore will not show in the box plot, and does not affect the distribution
df_intensity_log10 <-log10(df_intensity)

data_matrix_log10_melt<-reshape2::melt(as.matrix(df_intensity_log10))

colnames(data_matrix_log10_melt) <- c("Proteins", "Samples", "Log10(Inensity)")

p<-ggplot(data_matrix_log10_melt)+
    geom_boxplot(aes(x = Samples, y = `Log10(Inensity)`)) +
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  coord_flip()
  

plotly::ggplotly(p)

rm(p)
rm(data_matrix_log10_melt)


```




## Unique peptide number

<details>
  <summary>About this plot</summary>
  
  * The more unique peptide per protein, the better quality 

</details>

```{r distribution, eval =  TRUE, echo=FALSE,fig.width= 8,fig.height=6, error=TRUE}

peptide_summary <- as.data.frame(table(data_table_tidied$n_unique_peptides))
colnames(peptide_summary) = c("Unique_Peptides", "Freq")

p<- ggplot2::ggplot(data = peptide_summary)+
 geom_col(aes(x = Unique_Peptides,y = Freq))+
 theme_bw()
plotly::ggplotly(p)

```



## Protein Score distribution

<details>
  <summary>About this plot</summary>
  
  * The higher score the better quality
  * Different search engine usually have different score range

</details>


```{r echo=FALSE,fig.width= 8,fig.height=6, error=TRUE}

df_score <- data.frame(Score  = data_table_tidied$score)

p<- ggplot2::ggplot(data=df_score, aes(Score)) +
  geom_histogram(aes(y=..density..),color= "black",fill="white")+
  geom_density(alpha=.2, fill="#FF6666")+
  theme_bw()

plotly::ggplotly(p)

```


## Sparsity 

Columns starting with "LFQ_intensity_" will be selected if exist (meaning LFQ option was checked for label free quantification, which is deliberately processed by Maxquant already), otherwise Intensity_ columns will be used instead for protein experession.



```{r echo=FALSE,fig.width= 8,fig.height=6}

df_sparsity <- data_table_tidied$sparsity

p<- ggplot2::ggplot(data = df_sparsity)+
  geom_col(aes(x = sparsity,y = Freq))+
  xlab("Presence in the Protein Intensity Matrix")+
  theme_bw()
plotly::ggplotly(p)
```


<details>
  <summary> Figure shows the number of peptide in total with more than N presence, which helps to set the presence cutoff</summary>
</details>

```{r Sparsicty distribution plot, echo=FALSE,fig.width= 8,fig.height=6}
#df_sparsity_dec <- df_sparsity[order(df_sparsity$sparsity,decreasing = TRUE),]

df_sparsity$cumsum <-  cumsum(df_sparsity$Freq)

# redefine the sparsity sequence, in order to keep the order in the plot, This is the best way as far as I know
#df_sparsity_dec$sparsity <- factor(df_sparsity_dec$sparsity, levels = df_sparsity_dec$sparsity)

ggplot2::ggplot(data = df_sparsity)+
  geom_col(aes(x = sparsity,y = cumsum))+
  labs(title = "Sparsity Distribution", x = "Number of Presence",y = "Frequency") + 
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 90, hjust = 1),
        panel.grid = element_blank())



```






