---
title: "Test the rmarkdown template locally or remotely"
#output: html_notebook
---

# Readme

*  The following scripts show how to use r to upload a metalab result txt file to the opencpu server as the input, options with a meta file,  to render an html report then download the report to local storage
*  This function could be done by any script language, such as java, perl or bash
*  By doing this, Metalab can have the report easily, without the need to setup the r environment, but easily get all the report, especially with  Various data visualization in the report
*  These report are all html based and could be linked/iframed into a another frame file

* This doc shows the detailed method about how to test and debug the templates locally and remotely if this package is installed on remote server
* Current server is on compute Canada. The server is configured with a open cpu web server and open to the world.


* About how to make opencpu request is detailed here: https://www.opencpu.org/api.html



## Required Pacakges 

```{r}
#library(jsonlite)
#library(httr)

```


## preparation

* prepare the 5 result files from metalab from the same data set, and meta information table for this experiment, with the name of metainfo.txt.

see table format requirement for more information:
https://github.com/northomics/Metalab_development_wiki/wiki


* prepare the 5 markdown files for local test
* test locally first then remotely
* open the correstpoding rmarkdown file to debug

* setup the working directory as this file


# summary.txt


## local test

### without meta

```{r}
data_table <- rio::import("./extdata/summary.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)
# knit to all supported format (here we support pdf and html)
rmarkdown::render("./rmd/MQ_report_summary.Rmd",output_format = "html_document", params = list(data_table =  data_table), output_dir = getwd())




```



### with meta

```{r}
data_table <- rio::import("./extdata/summary.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)
meta_table <- rio::import("./extdata/metainfo.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE) 

rmarkdown::render("./rmd/MQ_report_summary.Rmd",output_format = "html_document", params = list(data_table =  data_table, meta_table = meta_table),  output_dir = getwd())

```



## remote test


```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_MQsummary_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")


```

### without meta

```{r}


# upload file and do the rendering
# in this case, the summary.txt is in the working dir. it can be anywhere with the path
# meta file is optional

r <- httr::POST(url_api, body = list(file = httr::upload_file("summary.txt")))
#r <- httr::POST(url_api, body = list(file = httr::upload_file("summary.txt"), meta = httr::upload_file("summary_metainfo.txt")))

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_ID_summary.html")



```

### with meta

```{r}


# upload file and do the rendering
# in this case, the summary.txt is in the working dir. it can be anywhere with the path
# meta file is optional

#r <- httr::POST(url_api, body = list(file = httr::upload_file("summary.txt")))
r <- httr::POST(url_api, body = list(file = httr::upload_file("summary.txt"), meta = httr::upload_file("metainfo.txt")))

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_ID_summary.html")


```






# proteingroups.txt

* proteinGroups_report.txt is a shorted and simplified version of proteinGroups.txt, much smaller in size, more suitable for report
* proteinGroups_report.txt does not have the correct number of some items, e.g. the number of contaminant etc



## local test

### without meta

```{r}
data_table  <- read.delim("proteinGroups_report.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE) # NOTE the read in options

rmarkdown::render("MQ_report_proteinGroups.Rmd",output_format = "html_document", params = list(input_datatable =  data_table), output_file="report_proteinGroups_summary.html")

```


### with meta

```{r}
data_table  <- read.delim("proteinGroups.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE) # NOTE the read in options
meta_table <- read.delim("metainfo.txt", header = TRUE, check.names = FALSE, stringsAsFactors = FALSE) # with meta file


rmarkdown::render("MQ_report_proteinGroups.Rmd",output_format = "html_document", params = list(input_datatable =  data_table, meta_table = meta_table), output_file="report_proteinGroups_summary.html")



```





##  remote test

### without meta

```{r}

url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_proteinGroups_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")


# upload file and do the rendering
# in this case, the proteinGroups.txt is in the working dir. it can be anywhere with the path
# variable r is the returning information from the curl function
r <- httr::POST(url_api, body = list(file = httr::upload_file("proteinGroups_report.txt")))
#r <- httr::POST(url_api, body = list(file = httr::upload_file("proteinGroups1.txt"), meta = httr::upload_file("proteinGroups1_meta.txt")), httr::timeout(200000))

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]
#paths

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_proteinGroups_summary.html")

```


### with meta

```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_proteinGroups_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")


# upload file and do the rendering
# in this case, the proteinGroups.txt is in the working dir. it can be anywhere with the path
# variable r is the returning information from the curl function
#r <- httr::POST(url_api, body = list(file = httr::upload_file("final_proteins.tsv")))
r <- httr::POST(url_api, body = list(file = httr::upload_file("proteinGroups_report.txt"), meta = httr::upload_file("metainfo.txt")), httr::timeout(200000))


# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]
paths


# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_proteinGroups_summary.html")


```




# peptide.txt test


## locally


### without meta

```{r}

# test on locoal drive, using local rmd
data_table <- rio::import("./extdata/peptides_report.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)

rmarkdown::render("./rmd/MQ_report_peptides.Rmd",output_format = "html_document", params = list(data_table =  data_table), output_dir = getwd())


```




### with meta

```{r}

data_table  <- rio::import("./extdata/peptides_report.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE) # NOTE the readin options
meta_table <- rio::import("./extdata/metainfo.txt", header = TRUE, check.names = FALSE, stringsAsFactors = FALSE) # with meta file

rmarkdown::render("./rmd/MQ_report_peptides.Rmd",output_format = "html_document", params = list(data_table =  data_table, meta_table = meta_table), output_file="MQ_report_peptides.html",output_dir = getwd())


```





## remotely


### without meta

```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_peptides_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# variable r is the returning information from the curl function
r <- httr::POST(url_api, body = list(file = httr::upload_file("peptides.txt")))
#r <- httr::POST(url_api, body = list(file = httr::upload_file("peptides3.txt"),meta = httr::upload_file("peptides3_meta.txt")), httr::timeout(200000))
r$status_code

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]
path_target
# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_peptides_summary.html")


```



### with meta

```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_peptides_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# in this case, the proteinGroups.txt is in the working dir. it can be anywhere with the path
# variable r is the returning information from the curl function
#r <- httr::POST(url_api, body = list(file = httr::upload_file("peptides.txt")))
r <- httr::POST(url_api, body = list(file = httr::upload_file("peptides.txt"),meta = httr::upload_file("metainfo.txt")), httr::timeout(200000))
r$status_code

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]
path_target
# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_peptides_summary.html")


```



# BuiltIn.taxa.refine.csv ----

## remotely

### without meta

```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_taxon_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# variable r is the returning information from the curl function
r <- httr::POST(url_api, body = list(file = httr::upload_file("BuiltIn.taxa.refine.csv")))

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_taxonomy_summary.html")


```


### with meta

```{r}
url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_taxon_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# variable r is the returning information from the curl function

r <- httr::POST(url_api, body = list(file = httr::upload_file("BuiltIn.taxa.refine.csv"),meta = httr::upload_file("metainfo.txt")), httr::timeout(200000))

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_taxonomy_summary.html")


```




# function.csv ----

## remote ----

### without meta

```{r}


url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_function_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# variable r is the returning information from the curl function
r <- httr::POST(url_api, body = list(file = httr::upload_file("extdata/functions.csv")))

r$status_code

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_function_summary.html")



```



### with meta

```{r}


url_api <- "http://206.12.91.148/ocpu/library/rmdocpu/R/render_function_file"

# get the root url
url_api_split <- strsplit(url_api, "/")[[1]]
url_server<- paste0(url_api_split[1],"//", url_api_split[3],"/")

# upload file and do the rendering
# variable r is the returning information from the curl function
#r <- httr::POST(url_api, body = list(file = httr::upload_file("functions.csv")))
r <- httr::POST(url_api, body = list(file = httr::upload_file("functions.csv"), meta = httr::upload_file("metainfo.txt")))

r$status_code

# get all the paths of all files from the opencpu end, and locate the one, which is the report
# this step needs to be done in the script enviroment

paths <- strsplit(rawToChar(r$content), "\n")[[1]]
path_target <- paths[grep("output.html",paths)]

# save/download the report file to local storage
# the file  "maxquant_result_summary.html" now is the report
curl::curl_download(paste0(url_server, path_target), "report_function_summary.html")



```


## locally

### without meta

```{r}


data_table <- read.delim("extdata/functions.csv", header = TRUE,sep= ",",check.names = TRUE, stringsAsFactors = FALSE)
#data_table <- read.delim("extdata/functions_20211015.csv", header = TRUE,sep= ",",check.names = TRUE, stringsAsFactors = FALSE)
# knit to all supported format (here we support pdf and html)

data_table <- read.delim("extdata/functions_20211018.tsv", header = TRUE,sep= "\t",quote="", check.names = TRUE, stringsAsFactors = FALSE,)



data_table <- rio::import("extdata/functions_20211018.tsv",check.names = TRUE)



rmarkdown::render("rmd/ML_report_function.Rmd",output_format = "html_document", params = list(input_datatable =  data_table))



```

### with meta

```{r}

### with meta

data_table <- read.delim("functions.csv", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)
meta_table <- read.delim("metainfo.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE) 

rmarkdown::render("ML_report_function.Rmd",output_format = "html_document", params = list(input_datatable =  data_table, meta_table = meta_table ))


```

# pepFunk 

## locally

```{r}

data_table  <- data_table <- read.delim("./extdata/peptides.txt", row.names = 1) # NOTE the readin options

meta_table <- rio::import("./extdata/metainfo_pepfunk.txt", header = TRUE, check.names = FALSE, stringsAsFactors = FALSE) # with meta file

#meta_table <- read.delim("./extdata/metadata.csv", header=F, sep=",")


rmarkdown::render("./rmd/ML_report_pepFunk.Rmd",output_format = "html_document", output_dir = getwd(), params = list(data_table =  data_table, meta_table = meta_table ))



rmarkdown::render("./rmd/ML_report_pepFunk.Rmd",output_format = "html_document", output_dir = getwd(), envir = new.env(), params = list(data_table =  data_table, meta_table = meta_table ))



# for debugging

#params$data_table <- data_table
#params$meta_table <- meta_table






```


